A Type System for Safe and Scalable Parallelism using Logical Regions \\ (Extended Version)

### 2012-07-16 doc/popl2013full/langdef.tex
proof

\typeenvx[G={\Gamma[id/T_1]}] e_2 : T_2

@@

\infrule[T-Let]{\typeenvx e_1 : T_1 \\ \typeenvx[G={\Gamma[id/T_1]}] e_2 : T_2}{\typeenvx : \text{let}\ id : T_1 \text{in}\ e_2 : T_2} &

@@

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
\mapconsist{\Omega} & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
\localconsist{L}{\Gamma} & \Leftrightarrow & \forall (id,v) \in L. \valueconsist[]{v}{M \llbracket \Gamma \rrbracket (id)} \\
\storeconsist{S}{H} & \Leftrightarrow & \forall (l,v) \in S. \valueconsist[]{v}{H(l)} \\
\end{array} & 
\begin{array}{lll}
\traceconsist{E} & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow \valueconsist[]{v}{H(l)})\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \Rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge \valueconsist[]{v}{\hat T_2}) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
\valueconsist[]{bv}{bool} \\
\valueconsist[]{iv}{int} \\
\valueconsist[]{null}{\hat T@\rho} \\
\end{array} &
\begin{array}{lll}
\valueconsist[]{l}{\hat T@\rho} & \Leftrightarrow & l \in \rho \wedge H(l) = \hat T \\
\valueconsist[]{\tuple{v_1}{v_2}}{\tuple{\hat T_1}{\hat T_2}} & \Leftrightarrow & (\valueconsist[]{v_1}{\hat T_1}) \wedge (\valueconsist[]{v_2}{\hat T_2}) \\
\valueconsist[]{\langle \langle \oton{\rho}, v \rangle \rangle}{\text{rr}[\oton{r}]\ \hat T\text{ where }\Omega} & \Leftrightarrow & (\valueconsist[]{v}{T[\rho_1/r_1, \ldots \rho_n/r_n]}) \wedge (\mapconsist[\{ (r_i, \rho_i) \}]{\Omega}) \\
\valueconsist[]{K}{coloring(\rho)} & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\vspace{-5mm}
\end{figure*}

@@

The proof is a standard induction on the structure of the derivation
For each of 
the Core Legion expressions there are six separate parts to the proof,
one for each of the three consistency properties that must be shown
for the execution environment of subexpressions and one for each of
the three properties of the execution's results.
Many of the cases are similar, and benefit from
the use of the following straighforward lemmas:

\begin{lem}
\label{lemma:heapconst:apply}
\rm
If $\storeconsist{S}{H}$ and $\traceconsist{E}$, then $\storeconsist{\text{apply}(S, E)}{H}$.
\end{lem}

@@

\begin{proof}
The proof is by cases on the structure of the derivation of $e$.
\begin{itemize}

\item
Case [E-Bool]. By assumption, 
$\opsenv{bv}{bv}{[]}$ and
$\typeenv{bv}{T}$ and
$\mapconsist{\Omega}$ and
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply is [T-Bool], so $T = bool$.
Therefore $\valueconsist{bv}{bool}$ and 
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$ all follow trivially.

\item Case [E-Int].  
By assumption, 
$\opsenv{iv}{iv}{[]}$ and 
$\typeenv{iv}{T}$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Int], so $T = int$.
Therefore $\valueconsist{iv}{int}$ and 
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$ all follow trivially.

\item Case [E-Null].
By assumption, 
$\opsenv{\text{null\ }T@r}{null}{[]}$ and 
$\typeenv{\text{null}}{T'}$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Null], so $T' = T@r$.
Therefore $\valueconsist{null}{T@r}$, $\traceconsist{[]}$ and $\privconsist{[]}{\Phi}$ all follow trivially.

\item
Case [E-MakeTuple].
By assumption, 
$\opsenv{\langle e_1, e_2 \rangle}{\langle v_1,v_2 \rangle}{E'}$ and 
$\typeenv{\langle e_1, e_2 \rangle}{T}$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-MakeType], so $T = \langle T_1, T_2 \rangle$.

For the subexpression $e_1$, 
from the form of rule [E-MakeTuple], we know that $\opsenv{e_1}{v_1}{E_1}$.
From the form of rule [T-MakeTuple], we know that $\typeenv{e_1}{T_1}$.
Then, by induction, we have $\valueconsist{v_1}{T_1}$, $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from the form of rule [E-MakeTuple], we know that $\opsenv[S=S']{e_2}{v_2}{E_2}$
where $S' = apply(S,E_1)$.  By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S' = \text{apply}(S,E)}{H}$.
From the form of rule [T-MakeTuple], we know that $\typeenv{e_2}{T_2}$.
Then, by induction, we have $\valueconsist{v_2}{T_2}$, $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

Finally, from the precedinig steps we can conclude that $\valueconsist{\langle v_1, v_2\rangle}{\langle T_1, T_2 \rangle}$.
By Lemma~\ref{lemma:heapconst:effects1}, $\traceconsist{E_1 \dplus E_2}$, and by Lemma~\ref{lemma:effsound:effects1}
$\privconsist{E_1 \dplus E_2}{\Phi}$.

\item Case [E-Tuple1].
By assumption, 
$\opsenv{e\text{.1}}{v_1}{E}$ and 
$\typeenv{e\text{.1}}{T_1}$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
From the form of [E-Tuple1] we know that
$\opsenv{e}{\langle v_1, v_2 \rangle}{E}$.
The only type rule that can apply is [T-Tuple1], and from the form of the rule we know
$\typeenv{e}{\langle T_1, T_2 \rangle}$.
By induction, we have $\valueconsist{\langle v_1, v_2 \rangle}{\langle T_1, T_2 \rangle}$, $\traceconsist{E}$ and 
      $\privconsist{E}{\Phi}$.
It follows immediately that $\valueconsist{v_1}{T_1}$ also holds.

\item Case[E-New].
By assumption, 
$\opsenv{\text{new }T@r}{l}{[]}$ and 
$\typeenv{\text{new }T@r}{T@r}$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The hypothesis of the [E-New] rule is that $l \in M(r)$ and $H(l) = M \llbracket T \rrbracket$,
which establishes $\valueconsist{l}{T}$.  
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.
\item Case [E-Var].
By assumption, 
$\opsenv{id}{v}{[]}$ and 
$\typeenv{id}{T}$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The hypothesis of the [E-Var] rule is $L(id) = v$.
From $\localconsist{L}{\Gamma}$ we have $\valueconsist{L(id)}{\Gamma(id)}$.
The hypothesis of the [T-Var] rule is $\Gamma(id) = T$.
Putting these three facts together we conclude $\valueconsist{v}{T}$.
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.

\item Case [E-IsNull-F].
By assumption, 
$\opsenv{\text{isnull}(e)}{\textit{false}}{E}$ and 
$\typeenv{\text{isnull}(e)}{bool}$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
From the form of [E-IsNull-F] we know that
$\opsenv{e}{l}{E}$, and
and from the form of the type rule [T-IsNull] we know
$\typeenv{e}{T@(\oton{r})}$.
Clearly $\valueconsist{\textit{false}}{bool}$, and
$\traceconsist{E}$ and 
      $\privconsist{E}{\Phi}$ follow by induction.

\item Case [E-IsNull-T] is almost identical to case [E-IsNull-F], with $l$ replaced by \textit{null} and
\textit{true} replaced by \textit{false}.

\item[E-Let]
\item[E-Add]
\item[E-Compare]
\item[E-IfElse-T]
\item[E-IfElse-F]
\item[E-Newcolor]
\item[E-Color]
\item[E-UpRgn]
\item[E-DownRgn]
\item[E-Read]
\item[E-Write]
\item[E-Reduce]
\item[E-Pack]
\item[E-Unpack]
\item[E-Call]
\item[E-Program]

\end{itemize}

\end{proof}
\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\mapconsist{\Omega}$ - Three expressions have subexpressions that modify $M$ or $\Omega$ and therefore do not
trivially satisfy region mapping consistency.  For {\tt partition},
the consistency of the coloring preserves region mapping consistency
with respect to the constraints.  For {\tt unpack}, the consistency of
a region relation instance guarantees consistency of region mapping.
Finally, the body of a called function uses an initially-empty set of
constraints, which are trivially satisfied.

@@

 \begin{table*}
 \centering
 {\small
 \begin{tabular}{l|p{5cm}|p{5cm}|p{5cm}}
 {\bf Rule} & 
 \multicolumn{1}{|c|}{$\mathbf{v \sim_H M \llbracket T \rrbracket}$} &
 \multicolumn{1}{|c|}{$\mathbf{E \sim H}$} &
 \multicolumn{1}{|c|}{$\mathbf{E : \Phi}$} \\
 \hline
 BoolConst & & & \\
 IntConst & & & \\
 TupleBuild & & & \\
 TupleField & & & \\
 LocalVar & & & \\
 NullPtr & & & \\
 NewPtr & & & \\
 IsNull & & & \\
 UpRegion & & & \\
 DownRegion & & & \\
 Read & & & \\
 Write & & & \\
 Reduce & & & \\
 Newcolor & & & \\
 Color & & & \\
 IntMath & & & \\
 IntCmp & & & \\
 Let & & & \\
 IfElse & & & \\
 TaskCall & & & \\
 Partition & & & \\
 Pack & & & \\
 Unpack & & &
 \end{tabular}}
 \caption{Proof pieces}
 \end{table*}

>>>>

\typeenvx[G={\Gamma[T_1/id]}] e_2 : T_2

@@

\infrule[T-Let]{\typeenvx e_1 : T_1 \\ \typeenvx[G={\Gamma[T_1/id]}] e_2 : T_2}{\typeenvx : \text{let}\ id : T_1 \text{in}\ e_2 : T_2} &

@@

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
\mapconsist{\Omega} & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
\localconsist{L}{\Gamma} & \Leftrightarrow & \forall (id,v) \in L. \valueconsist[]{v}{M \llbracket \Gamma \rrbracket (id)} \\
\storeconsist{S}{H} & \Leftrightarrow & \forall (l,v) \in S. \valueconsist[]{v}{H(l)} \\
\end{array} & 
\begin{array}{lll}
\traceconsist{E} & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow \valueconsist[]{v}{H(l)})\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge \valueconsist[]{v}{\hat T_2}) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
\valueconsist[]{bv}{bool} \\
\valueconsist[]{iv}{int} \\
\valueconsist[]{null}{\hat T@\rho} \\
\end{array} &
\begin{array}{lll}
\valueconsist[]{l}{\hat T@\rho} & \Leftrightarrow & l \in \rho \wedge H(l) = \hat T \\
\valueconsist[]{\tuple{v_1}{v_2}}{\tuple{\hat T_1}{\hat T_2}} & \Leftrightarrow & (\valueconsist[]{v_1}{\hat T_1}) \wedge (\valueconsist[]{v_2}{\hat T_2}) \\
\valueconsist[]{\langle \langle \oton{\rho}, v \rangle \rangle}{\text{rr}[\oton{r}]\ \hat T\text{ where }\Omega} & \Leftrightarrow & (\valueconsist[]{v}{T[\rho_1/r_1, \ldots \rho_n/r_n]}) \wedge (\mapconsist[\{ (r_i, \rho_i) \}]{\Omega}) \\
\valueconsist[]{K}{coloring(\rho)} & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\vspace{-5mm}
\end{figure*}

@@

The proof is a standard induction on the structure of the derivation
For each of 
the Core Legion expressions there are six separate parts to the proof,
one for each of the three consistency properties that must be shown
for the execution environment of subexpressions and one for each of
the three properties of the execution's results.
Many of the cases are similar, and benefit from
the use of the following straighforward lemmas:
\begin{lem}
\label{lemma:heapconst:apply}
\rm
If $\storeconsist{S}{H}$ and $\traceconsist{E}$, then $\storeconsist{\text{apply}(S, E)}{H}$.
\end{lem}

@@

We first give an outline of the idea behindthe proof.  For each subexpression of a compound expression,
there are three consistency properties we must prove to be able to apply induction: \\

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\mapconsist{\Omega}$ - Three expressions have subexpressions that modify $M$ or $\Omega$ and therefore do not
trivially satisfy region mapping consistency.  For {\tt partition},
the consistency of the coloring preserves region mapping consistency
with respect to the constraints.  For {\tt unpack}, the consistency of
a region relation instance guarantees consistency of region mapping.
Finally, the body of a called function uses an initially-empty set of
constraints, which are trivially satisfied.

@@

\begin{proof}
The proof of the theorem is by cases on the structure of the derivation of $e$.
\begin{itemize}

\item
Case [E-Bool]. By assumption, 
$$\opsenv{bv}{bv}{[]}$$ and
$$\typeenv{bv}{T}$$ and
$\mapconsist{\Omega}$ and
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply is [T-Bool], so $T = bool$.
Therefore $\valueconsist{bv}{bool}$ and 
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$ all follow trivially.

\item Case [E-Int].  
By assumption, 
$$\opsenv{iv}{iv}{[]}$$ and 
$$\typeenv{iv}{T}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Int], so $T = int$.
Therefore $\valueconsist{iv}{int}$ and 
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$ all follow trivially.

\item Case [E-Null].
By assumption, 
$$\opsenv{\text{null\ }T@r}{null}{[]}$$ and 
$$\typeenv{\text{null}}{T'}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Null], so $T' = T@r$.
Therefore $\valueconsist{null}{T@r}$, $\traceconsist{[]}$ and $\privconsist{[]}{\Phi}$ all follow trivially.

\item
Case [E-MakeTuple].
By assumption, 
$$\opsenv{\langle e_1, e_2 \rangle}{\langle v_1,v_2 \rangle}{E'}$$ and 
$$\typeenv{\langle e_1, e_2 \rangle}{T}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-MakeTuple], so $T = \langle T_1, T_2 \rangle$.

For the subexpression $e_1$, 
from the form of rule [E-MakeTuple], we know that $$\opsenv{e_1}{v_1}{E_1}$$
From the form of rule [T-MakeTuple], we know that $$\typeenv{e_1}{T_1}$$
Then, by induction, we have $\valueconsist{v_1}{T_1}$, $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from the form of rule [E-MakeTuple], we know that $$\opsenv[S=S']{e_2}{v_2}{E_2}$$
where $S' = apply(S,E_1)$.  By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S' = \text{apply}(S,E)}{H}$.
From the form of rule [T-MakeTuple], we know that $$\typeenv{e_2}{T_2}$$
Then, by induction, we have $\valueconsist{v_2}{T_2}$, $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

From the preceding steps we can conclude that $\valueconsist{\langle v_1, v_2\rangle}{\langle T_1, T_2 \rangle}$.
Finally, from [E-MakeTuple] we also have $$valid\_interleave(S,C,E',E_1,E_2)$$ and
Lemma~\ref{lemma:heapconst:effects2}, $\traceconsist{E'}$ and by Lemma~\ref{lemma:effsound:effects2}
$\privconsist{E'}{\Phi}$.

\item Case [E-Tuple1].
By assumption, 
$$\opsenv{e\text{.1}}{v_1}{E}$$ and 
$$\typeenv{e\text{.1}}{T_1}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
From the form of [E-Tuple1] we know that
$$\opsenv{e}{\langle v_1, v_2 \rangle}{E}$$
The only type rule that can apply is [T-Tuple1], and from the form of the rule we know
$$\typeenv{e}{\langle T_1, T_2 \rangle}$$
By induction, we have $\valueconsist{\langle v_1, v_2 \rangle}{\langle T_1, T_2 \rangle}$, $\traceconsist{E}$ and 
      $\privconsist{E}{\Phi}$.
It follows immediately that $\valueconsist{v_1}{T_1}$ also holds.

\item Case[E-New].
By assumption, 
$$\opsenv{\text{new }T@r}{l}{[]}$$ and 
$$\typeenv{\text{new }T@r}{T@r}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The hypothesis of the [E-New] rule is that $l \in M(r)$ and $H(l) = M \llbracket T \rrbracket$,
which establishes $\valueconsist{l}{T}$.  
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.
\item Case [E-Var].
By assumption, 
$$\opsenv{id}{v}{[]}$$ and 
$$\typeenv{id}{T}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The hypothesis of the [E-Var] rule is $L(id) = v$.
From $\localconsist{L}{\Gamma}$ we have $\valueconsist{L(id)}{\Gamma(id)}$.
The hypothesis of the [T-Var] rule is $\Gamma(id) = T$.
Putting these three facts together we conclude $\valueconsist{v}{T}$.
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.

\item Case [E-IsNull-F].
By assumption, 
$$\opsenv{\text{isnull}(e)}{\textit{false}}{E}$$ and 
$$\typeenv{\text{isnull}(e)}{bool}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
From the form of [E-IsNull-F] we know that
$\opsenv{e}{l}{E}$, and
and from the form of the type rule [T-IsNull] we know
$\typeenv{e}{T@(\oton{r})}$.
Clearly $\valueconsist{\textit{false}}{bool}$, and
$\traceconsist{E}$ and 
      $\privconsist{E}{\Phi}$ follow by induction.

\item Case [E-IsNull-T] is almost identical to case [E-IsNull-F], with $l$ replaced by \textit{null} and
\textit{false} replaced by \textit{true}.

\item Case [E-Let].
By assumption, 
$$\opsenv{\text{let } id:T_1 = e_1 \text{ in } e_2}{v_2}{E'}$$ and 
$$\typeenv{\text{let } id:T_1 = e_1 \text{ in } e_2}{T_2}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.

For the subexpression $e_1$, 
from [E-Let], we know that $$\opsenv{e_1}{v_1}{E_1}$$
From type rule [T-Let], we know that $$\typeenv{e_1}{T_1}$$
Then, by induction, we have $\valueconsist{v_1}{T_1}$, $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from [E-Let], we have $$\opsenv[L=L',S=S']{e_2}{v_2}{E_2}$$
where $L' = L[v_1/id]$ and $S' = apply(S,E_1)$.  
From [T-Let], we have $\typeenv[G={\Gamma[T_1/id]}]{e_2}{T_2}$.
By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S' = \text{apply}(S,E)}{H}$.
From $\localconsist{L}{\Gamma}$ and $\valueconsist{v_1}{T_1}$ we conclude $\localconsist{L[v_1/id]}{\Gamma[T_1/id]}$.
Then, by induction, we have $\valueconsist{v_2}{T_2}$, $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

Finally, from [E-Let] we also have $$valid\_interleave(S,C,E',E_1,E_2)$$
Then, by Lemma~\ref{lemma:heapconst:effects2}, $\traceconsist{E'}$ and by 
Lemma~\ref{lemma:effsound:effects2} $\privconsist{E'}{\Phi}$.

\item Case [E-Add].
By assumption, 
$$\opsenv{e_1 + e_2}{v'}{E'}$$ and 
$$\typeenv{e_1 + e_2}{T}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Add], so $T = int$.

For the subexpression $e_1$, 
from [E-Add] we know $$\opsenv{e_1}{v_1}{E_1}$$
From [T-Add], we know that $$\typeenv{e_1}{int}$$
Then, by induction, we have $\valueconsist{v_1}{T_1}$, $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from [E-Add], we know $$\opsenv[S=S']{e_2}{v_2}{E_2}$$
where $S' = apply(S,E_1)$.  By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S' = \text{apply}(S,E)}{H}$.
From [T-Add] we know $$\typeenv{e_2}{int}$$
Then, by induction, we have $\valueconsist{v_2}{T_2}$, $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

It is immediate that $\valueconsist{v_1 + v_2}{int}$.
Finally, from [E-Add] we also have $$valid\_interleave(S,C,E',E_1,E_2)$$ and
Lemma~\ref{lemma:heapconst:effects2}, $\traceconsist{E'}$ and by Lemma~\ref{lemma:effsound:effects2},
$\privconsist{E'}{\Phi}$.

\item Case [E-Compare] is isomorphic to [E-Add].

\item Case [E-IfElse-T].
 By assumption, 
$$\opsenv{\text{ if } e_1 \text{ then } e_2 \text{ else } e_3}{v_2}{E}$$ and 
$$\typeenv{\text{ if } e_1 \text{ then } e_2 \text{ else } e_3}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

For the subexpression $e_1$, 
from [E-IfElse-T] we know $$\opsenv{e_1}{true}{E_1 \dplus E_2}$$
From [T-IfElse], we know that $$\typeenv{e_1}{bool}$$
Clearly $\valueconsist{true}{bool}$, and by induction we have  $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from [E-IfElse-T] we know $$\opsenv[S=S']{e_2}{v_2}{E_2}$$
where $S' = apply(S,E_1)$.  By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S' = \text{apply}(S,E)}{H}$.
From [T-IfElse-T] we know $$\typeenv{e_2}{T}$$
Then, by induction, we have $\valueconsist{v_2}{T}$, $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

Finally, by Lemma~\ref{lemma:heapconst:effects1}, $\traceconsist{E_1 \dplus E_2}$ and by Lemma~\ref{lemma:effsound:effects1},
$\privconsist{E_1 \dplus E_2}{\Phi}$.

\item Case [E-IfElse-F] is nearly the same as [E-IfElse-T] with {\em false} replacing {\em true} and $e_3$ replacing $e_2$.

\item Case [E-Newcolor].
By assumption, 
$$\opsenv{\text{newcolor } r}{K}{[]}$$ 
where  $K = \{ (l_1,v_1),\ldots,(l_p,i_p) \}$ such that
$l_i \in M(r)$ for all $i$ and $l_1,\ldots,l_p$ are pairwise distinct locations.
In addition, we have the assumptions
$$\typeenv{\text{newcolor } r}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply if [T-Newcolor], so can infer that $T = \text{coloring}(r)$.
From the structure of $K$ it follows that $\valueconsist{K}{\text{coloring(r)}}$.
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.

\item  Case [E-Color].
By assumption, 
$$\opsenv{\text{color($e_1$, $e_2$, $e_3$)}}{K'}{E'}$$ and
$$\typeenv{\text{color($e_1$, $e_2$,$e_3$)}}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply if [T-Color], so we know $T = \text{coloring}(r)$.

For the subexpression $e_1$, 
from [E-Color] we know $$\opsenv{e_1}{K}{E_1}$$
From [T-Color], we know that $$\typeenv{e_1}{\text{coloring}(r)}$$
By induction it follows that $\valueconsist{K}{\text{coloring(r)}}$ and $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from [E-Color] we know $$\opsenv[S=S']{e_2}{l}{E_2}$$ where $S' = \text{apply}(S,E_1)$.
From [T-Color], we know that $$\typeenv{e_2}{T'@r}$$
By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S'}{H}$.
By induction, $\valueconsist{l}{T'@r}$ and $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

For the subexpression $e_3$, 
from [E-Color] we know $$\opsenv[S=S'']{e_3}{v}{E_3}$$ where $S'' \text{apply}(S',E_2)$.
From [T-Color], we know that $$\typeenv{e_3}{int}$$
By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S''}{H}$.
By induction, $\valueconsist{v}{int}$ and $\traceconsist{E_3}$ and 
      $\privconsist{E_3}{\Phi}$.

Now $K'$ is built from $K$ by extending $K$ with a mapping for location $l$.  First, because $\valueconsist{l}{T'@r}$ and
$\valueconsist{K}{\text{coloring(r))}$ and because $K'$ excludes any previous mapping for $l$ (guaranteeing the the coloring is a function), we conclude $\valueconsist{K'}{\text{coloring}(r)}}$.  From [E-Color], we also know
$$valid\_interleave(S, C, E', E_1, E_2, E_3)$$ and by applying Lemma~\ref{lemma:heapconst:effects2} twice
we can conclude $\traceconsist{E'}$.  Similarly, by applying Lemma~\ref{lemma:effsound:effects2} twice
we can conlude $\privconsist{E'}{\Phi}$.

\item  Case [E-UpRgn].
By assumption, $$\opsenv{\text{upregion}(e,\oton{r})}{v}{E}$$ and
$$\typeenv{\text{upgregion}(e,\oton{r})}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The type rule must be [T-UpRgn], so we also know $T = T' @ (\oton{r})$.

From [E-UpRgn] we know $$\opsenv{e}{v}{E}$$ and
from [T-UpRgn] we know $$\typeenv{e}{T' @ (r_1',\ldots,r_k')}$$
By induction, we can conclude that $\valueconsist{v}{T'@(r_1',\ldots,r_k')}$,
$\traceconsist{E}$ and $\privconsist{E}{\Phi}$.

To show $\valueconsist{v}{T'@(\oton{r})}$, we first observe that from
$\valueconsist{v}{T'@(r_1',\ldots,r_k')}$ we know
there is an $i$ such that $v \in M(r_i')$ and 
$H(v) = \hat T'$.   From [T-UpRgn]
we know $\forall i. \exists j. r_i' \leq r_j \in \Omega^*$.  Combining this with
$\mapconsist{\Omega}$ and Lemma~\ref{lemma:omegaclosure}, we know there
is some $r_j$ such that $v \in M(r_j)$.
Therefore $\valueconsist{v}{T'@(\oton{r})}$.

\item Case [E-DownRgn].
By assumption, $$\opsenv{\text{downregion}(e,\oton{r})}{v}{E}$$ and
$$\typeenv{\text{downgregion}(e,\oton{r})}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The type rule must be [T-DownRgn], so we also know $T = T' @ (\oton{r})$.

From [E-DownRgn] we know $$\opsenv{e}{l}{E}$$ and
from [T-DownRgn] we know $$\typeenv{e}{T' @ (r_1',\ldots,r_k')}$$
By induction, we can conclude that $\valueconsist{l}{T'@(r_1',\ldots,r_k')}$,
$\traceconsist{E}$ and $\privconsist{E}{\Phi}$.

To show $\valueconsist{v}{T'@(\oton{r})}$ there are two cases.
If $v = null$ then $\valueconsist{null}{T'@(\oton{r})}$ follows by definition.
Otherwise, $v = l$ and $l \in M(r_i)$ for some $i$.  We also know $H(v) = \hat T'$
because $\valueconsist{l}{T'@(r_1',\ldots,r_k')}$.  Therefore
$\valueconsist{v}{T'@(\oton{r})}$.

\item Case [E-Read].
By assumption, $\opsenv{\text{read}(e)}{v}{E \dplus [ read(l,excl,v) ]}$ and
$\typeenv{\text{read}(e)}{T}$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

From [E-Read] we have $$\opsenv{e}{l}{E}$$ and
from [T-Read] we know $$\typeenv{e}{T @ (\oton{r})}$$
By induction, we can conclude that $\valueconsist{l}{T@(\oton{r})}$,
$\traceconsist{E}$ and $\privconsist{E}{\Phi}$.

To show $\valueconsist{v}{T}$ there are two cases.
If $v = M(l)$ then the claim follows from $\valueconsist{l}{T@(\oton{r})}$.
If $v$ is instead some $v'$ of type $H(l)$ the result also follows because $H(l) = \hat{T}$.
Next, $\traceconsist{E \dplus [ read(l,excl,v)]}$ follows immediately from $\traceconsist{E}$ as
any read operation is consistent (the resriction is only on writes and reductions).

Finally, we must show $\privconsist{E \dplus [ read(l,excl,v)]}{\Phi}$.
Since we have already shown $\privconsist{E}{\Phi}$, it suffices to show $\privconsist{[ read(l,excl,v) ]}{\Phi}$.
By $\valueconsist{l}{T@(\oton{r})}$, there is some $r_i$ such that $l \in M(r_i)$.  Furthermore,
from [T-Read] we have $reads(r_i) \in \Phi^*$, which shows $\privconsist{[ read(l,excl,v)]}{\Phi^*}$.
The result follows from Lemma~\ref{lemma:phiclosure}.
